---
title: "farsdata-tutorial"
author: "Rahul Venugopal"
date: "4th of May, 2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{farsdata-tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gradedassignment)
```
This package is a submission towards final assessment of `Building an R package` course.
This has two functions to obtain a summary and a visual map of count of fatal injuries in motor vehicle traffic crashes from 2013,2014, and 2015.

### Data
Three data files were provided by the course instructors. These are from the National Highway Traffic Safety Administration (NHTSA) Fatality Analysis Reporting System (FARS) data.

You may download the full data set from this [link](https://www.nhtsa.gov/nhtsa-ftp?fid=66#block-nhtsa-page-title)


```{r 2015_data}
filename <- make_filename('2015')
data_2015 <- fars_read(filename) 
knitr::kable(head(data_2015))

```

### Functions in the package

**1. `make_filename` function will fetch the filename**
```{r make_filename}
filename <- make_filename(2015)
filename
```

**2. `fars_read` function uses 'make_filename' and loads data**
```{r fars_read}
data <- fars_read(make_filename(2015) )
knitr::kable(head(data))
```

**3. `fars_read_years` function subsets data based on year**
```{r fars_read_years}
data <- fars_read_years(2015)
data
```

**4. `fars_summarize_years` generates a count summary of fatality numbers per month for years specified as output**
```{r fars_summarize_years}
summary <- fars_summarize_years(2013:2015)
summary
```

```{r fars_summarize_years single}
summary <- fars_summarize_years(2013)
summary
```

**5. `fars_map_state` function plots a map with dots representing location of each crime committed**
This function expects state ID number as first argument and year as second argument
```{r fars_map_state}
fars_map_state(49,2015)
```

```{r fars_map_state another}
fars_map_state(4,2014)
```


